Theories of deliberated choice

### 2024-01-23 simple.tex
Thm promotion
Notation hist
Attempts concrete not HAI

    For \cref{it:HAI} $\implies$ \cref{it:condHAI},
    pick an indicative claim $C$ and consider $i \in I$ and $\theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub$ such that $\pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$.
    Define $H = \theta^{-1}(i, P)$.
  \end{proof}
  \begin{corollary}
    \label{th:suffIndicDec}
    Assume that for some $p \in P$, we have $\exists i \in I \suchthat P_i = p$ and $\forall i_2 \suchthat P_{i_2} \neq p$ we have that $\exists A \in \finite(\allargs) \suchthat \forall i \in I \suchthat P_i \neq p, \exists a \in A \suchthat {\ileadstost}(a) \in P \setminus \set{p}$, then there exists a claim that holds and is anonymous and indicative.
  \end{corollary}
  \begin{proof}
    We know that $i_1$ has some decisive argument $a_1$ in favor of $p$.
    Assume (wlog) that $A = \set{2, …, \card{A} + 1}$.
    Define $\hist_1 = \emptyset$ and for $i \in \intvl{2, \card{A} + 1}$, $\hist_i = (\hist_{i - 1}, a_1, i)$.
    Define $H = \set{\hist_i \suchthat i \in \intvl{2, \card{A} + 1}}$.
    Apply \cref{th:suffIndic}.
  \end{proof}
\begin{example}
  Consider $I_p$ such that $\forall i \in I_p$: $i$ constantly says $p$.
  Consider $J = I \setminus I_p$ whose DJ is $p' \neq p$, and such that each $j \in J$ has a single decisive argument $\ar_j$ for $p'$, and when $\ar_j$ is not in $\hist$, $\hist \ileadsto[j] p$.
  Then, the condition is not satisfied.
\end{example}
\begin{conjecture}
  \label{th:caractNotIndic}
  The following conditions on $\allleadsto$ are equivalent:
  \begin{enumerate}[label=T\ref{th:caractNotIndic}-{\arabic*}]
    \item \label{it:NotHAI} there exists no claim that holds and is anonymous and indicative;
    \item \label{it:condNotHAI} all paths are such that for all $i_1$ having a DJ, there is an $i_2$ with another deliberated judgement that behaves identically on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I \suchthat \pi(i_1) \neq \emptyset: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)$);
    \item \label{it:condNotHAIInf} all paths are such that for all $i_1$, either at least one individual without a DJ, or infinitely many individuals $i_2$ with another deliberated judgement than $i_1$, behave identically to $i_1$ on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I: [\exists i_2 \in I \suchthat \pi(i_2) = \emptyset \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)] \lor [\card(\set{i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}) = \infty]$)
  \end{enumerate}
\end{conjecture}
\begin{proof}
  Let us show that NOT \ref{it:condNotHAI} implies \ref{it:condHAI}.
  Consider: some individual $i$ and some path $\hist$ are such that $\pi(i) \neq \emptyset$ and
  $\forall i_2 \in I \suchthat \pi(i) \neq \pi(i_2): \exists \histp \wsub \hist \suchthat {\ileadsto[i]}(\histp) \neq {\ileadsto[i_2]}(\histp)$.
  Define $H = \set{\histp \wsub \hist}$ and $i_1 = i$.

@@

  Let us now prove that \ref{it:condNotHAI} implies \ref{it:condNotHAIInf}.
  Consider any $\hist \in \allhists, i_1 \in I$.
  Define $J = \set{i_2 \in I \suchthat \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}$.
  Assume that $\forall i_2 \in J: \pi(i_2) \neq \emptyset$ (otherwise, the first disjunct of \ref{it:condNotHAIInf} holds), and let us show that $\card\set{i_2 \in J \suchthat \pi(i_1) \neq \pi(i_2)} = \infty$.
  From $\forall i_2 \in J: \pi(i_2) \neq \emptyset$, we obtain $\pi(i_1) \neq \emptyset$, thus, $\exists a_{i_1} \in \allargs \suchthat a_{i_1} \ileadstost[i_1] \pi(i_1)$.
  Apply \ref{it:condNotHAI} to obtain some $i_2 \in J$ such that $\pi(i_1) \neq \pi(i_2)$.
  Also, by our assumption, $\pi(i_2) \neq \emptyset$, thus, $\exists a_{i_2} \in \allargs \suchthat a_{i_2} \ileadstost[i_2] \pi(i_2)$.
  Consider the path $\histp = (\hist, a_{i_1}, a_{i_2})$. We have $\histp \ileadsto[i_1] \pi(i_1) \neq \histp \ileadsto[i_2] \pi(i_2)$, thus ${\ileadsto[i_1]}(\histp) \neq {\ileadsto[i_2]}(\histp)$. 
  Apply \ref{it:condNotHAI} on $\histp, i_1$. We obtain some $i_3 \in I$ such that $\pi(i_1) \neq \pi(i_3)$ and $\forall \histpp \wsub \histp: {\ileadsto[i_1]}(\histpp) = {\ileadsto[i_3]}(\histpp)$.
  In particular, ${\ileadsto[i_1]}(\histp) = {\ileadsto[i_3]}(\histp)$, thus, $i_3 \neq i_2$.
  Also, $i_3 \in J$.
  Iterating this process, we obtain an infinite sequence of distinct individuals in $J$.
  In other words, we have shown that if some $i_2 \in J, \hist^{(2)} \in \allhists$ with $\pi(i_2) \neq \emptyset$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(2)}$, then there exists $i_3 \in J, \hist^{(3)} \in \allhists$ with $\pi(i_3) \neq \emptyset$ and $\hist^{(2)} \ssub \hist^{(3)}$ and $i_1$ and $i_2$ disagree somewhere on $\hist^{(3)}$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(3)}$. (Should check this summary.)
  Could also use the following probably simpler proof. Consider any finite subset of $j$ individuals, build a $H'$ with all the test paths containing decisive arguments, namely $\hist^{(2)} = (\hist, a_1, a_2)$, $\hist^{(3)} = (\hist^{(2)}, a_1, a_3)$, …, and show that there’s another $j$ in fact.
\end{proof}
\begin{example}
  TODO find example where Per \cref{th:caractNotIndic}, there is no claim that holds and is anonymous and indicative.

@@

  The condition $\pi(i_1) \neq \emptyset$ in \ref{it:condNotHAI} is required, as seen by this example where \ref{it:condNotHAI} is satisfied, even though
  $[\forall \hist \in \allhists, i_1 \in I: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)]$ is not.
We could have \ref{it:condNotHAI} satisfied, but not the part with $\pi(i_1) \neq \emptyset$.
  Consider $i_1$ with empty DJ on some path has only individuals with identical behavior on the path that have an empty DJ (though for normal $i$’s, the condition is satisfied, thus there are collisions, meaning some individuals with different DJ having identical behavior on the path).
  TODO check.
\end{example}
\begin{remark}
  Given $J \wsub I$, say that a set of paths discriminates $J$ iff some function associating each path to an answer is compatible with some $j \in J$ (meaning that $f \wsub {\ileadsto[j]}$) and is compatible with no $i \notin J$.
  The condition says that there must be some $p$ and some increasing set of paths $H$ such that $H$ discriminates $J = \pi^{-1}(p)$.
\end{remark}
\begin{remark}
  For failing the condition, I need an infinity of $j$’s (and at least one $i$) who agree on any branch of the history tree? (Meaning relations $\ileadsto[j]$, we don’t care about individuals.)
  No: on any finite branch in the tree, I need that remains one $j$ and one $i$ such that infinite prolongation of the branch will never discriminate them.

>>>>

    For \cref{it:HAI} $\implies$ \cref{it:condHAI},
    pick an indicative claim $C$ and consider $i \in I$ and $\theta \in \Theta \cap (\allhists × \set{i} × P) \cap C^\wsub$ such that $\pi(\set{{\allleadstog} \suchthat \theta \wsub {\allleadstog}} \cap C)(i) \ssub P$.
    Define $H = \theta^{-1}(i, P)$.
  \end{proof}
  \begin{corollary}
    \label{th:suffIndicDec}
    Assume that for some $p \in P$, we have $\exists i \in I \suchthat P_i = p$ and $\forall i_2 \suchthat P_{i_2} \neq p$ we have that $\exists A \in \finite(\allargs) \suchthat \forall i \in I \suchthat P_i \neq p, \exists a \in A \suchthat {\ileadstost}(a) \in P \setminus \set{p}$, then there exists a claim that holds and is anonymous and indicative.
  \end{corollary}
  \begin{proof}
    We know that $i_1$ has some decisive argument $a_1$ in favor of $p$.
    Assume (wlog) that $A = \set{2, …, \card{A} + 1}$.
    Define $\hist_1 = \emptyset$ and for $i \in \intvl{2, \card{A} + 1}$, $\hist^{(i)} = (\hist_{i - 1}, a_1, i)$.
    Define $H = \set{\hist^{(i)} \suchthat i \in \intvl{2, \card{A} + 1}}$.
    Apply \cref{th:suffIndic}.
  \end{proof}
\begin{example}
  Consider $I_p$ such that $\forall i \in I_p$: $i$ constantly says $p$.
  Consider $J = I \setminus I_p$ whose DJ is $p' \neq p$, and such that each $j \in J$ has a single decisive argument $\ar_j$ for $p'$, and when $\ar_j$ is not in $\hist$, $\hist \ileadsto[j] p$.
  Then, the condition is not satisfied.
\end{example}
\begin{theorem}
  \label{th:caractNotIndic}
  The following conditions on $\allleadsto$ are equivalent:
  \begin{enumerate}[label=T\ref{th:caractNotIndic}-{\arabic*}]
    \item \label{it:NotHAI} there exists no claim that holds and is anonymous and indicative;
    \item \label{it:condNotHAI} all paths are such that for all $i_1$ having a DJ, there is an $i_2$ with another deliberated judgement that behaves identically on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I \suchthat \pi(i_1) \neq \emptyset: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)$);
    \item \label{it:condNotHAIInf} all paths are such that for all $i_1$, either at least one individual without a DJ, or infinitely many individuals $i_2$ with another deliberated judgement than $i_1$, behave identically to $i_1$ on the whole path (formally: $\forall \hist \in \allhists, i_1 \in I: [\exists i_2 \in I \suchthat \pi(i_2) = \emptyset \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)] \lor [\card(\set{i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}) = \infty]$)
  \end{enumerate}
\end{theorem}
\begin{proof}
  Let us show that NOT \ref{it:condNotHAI} implies \ref{it:condHAI}.
  Consider: some individual $i$ and some path $\hist$ are such that $\pi(i) \neq \emptyset$ and
  $\forall i_2 \in I \suchthat \pi(i) \neq \pi(i_2): \exists \histp \wsub \hist \suchthat {\ileadsto[i]}(\histp) \neq {\ileadsto[i_2]}(\histp)$.
  Define $H = \set{\histp \wsub \hist}$ and $i_1 = i$.

@@

  Let us now prove that \ref{it:condNotHAI} implies \ref{it:condNotHAIInf}.
  Consider any $\hist \in \allhists, i_1 \in I$.
  Define $J = \set{i_2 \in I \suchthat \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)}$.
  Assume that $\forall i_2 \in J: \pi(i_2) \neq \emptyset$ (otherwise, the first disjunct of \ref{it:condNotHAIInf} holds), and let us show that $\card\set{i_2 \in J \suchthat \pi(i_1) \neq \pi(i_2)} = \infty$.
  From $\forall i_2 \in J: \pi(i_2) \neq \emptyset$, we obtain $\pi(i_1) \neq \emptyset$, thus, $\exists a_{i_1} \in \allargs \suchthat a_{i_1} \ileadstost[i_1] \pi(i_1)$.
  Apply \ref{it:condNotHAI} to obtain some $i_2 \in J$ such that $\pi(i_1) \neq \pi(i_2)$.
  Also, by our assumption, $\pi(i_2) \neq \emptyset$, thus, $\exists a_{i_2} \in \allargs \suchthat a_{i_2} \ileadstost[i_2] \pi(i_2)$.
  Consider the path $\histp = (\hist, a_{i_1}, a_{i_2})$. We have $\histp \ileadsto[i_1] \pi(i_1) \neq \histp \ileadsto[i_2] \pi(i_2)$, thus ${\ileadsto[i_1]}(\histp) \neq {\ileadsto[i_2]}(\histp)$. 
  Apply \ref{it:condNotHAI} on $\histp, i_1$. We obtain some $i_3 \in I$ such that $\pi(i_1) \neq \pi(i_3)$ and $\forall \histpp \wsub \histp: {\ileadsto[i_1]}(\histpp) = {\ileadsto[i_3]}(\histpp)$.
  In particular, ${\ileadsto[i_1]}(\histp) = {\ileadsto[i_3]}(\histp)$, thus, $i_3 \neq i_2$.
  Also, $i_3 \in J$.
  Iterating this process, we obtain an infinite sequence of distinct individuals in $J$.
  In other words, we have shown that if some $i_2 \in J, \hist^{(2)} \in \allhists$ with $\pi(i_2) \neq \emptyset$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(2)}$, then there exists $i_3 \in J, \hist^{(3)} \in \allhists$ with $\pi(i_3) \neq \emptyset$ and $\hist^{(2)} \ssub \hist^{(3)}$ and $i_1$ and $i_2$ disagree somewhere on $\hist^{(3)}$ and $i_1$ and $i_2$ agree everywhere on $\hist^{(3)}$. (Should check this summary.)
  Could also use the following probably simpler proof. Consider any finite subset of $j$ individuals, build a $H'$ with all the test paths containing decisive arguments, namely $\hist^{(2)} = (\hist, a_1, a_2)$, $\hist^{(3)} = (\hist^{(2)}, a_1, a_3)$, …, and show that there’s another $j$ in fact.
\end{proof}
\begin{example}
  Given $\hist \in \allhists$, $\ar \in \allargs$ and $p \in P$, define ${\ileadsto[i^{(\hist, p)}]}$ as $\histp \ileadsto[i^{(\hist, p)}] \phi$ if $\histp \wsub \hist$ and $\histp \ileadsto[i^{(\hist, p)}] p$ otherwise.
  It follows that $\forall \ar \notin \hist(\N)$, then $\ar \ileadstost[i^{(\hist, \ar)}] p$.

  Given any infinite set $\allargs$, define $I = \set{i^{(\hist, p)} \suchthat \hist \in \allhists, p \in P}$.
  Considering any $\hist \in \finite(\allhists)$ and $i^{(\histp, p)} \in I$, pick any $p' \neq p$ and consider $j = i^{(\histp, p')}$.

  NOPE

  Given $H_\phi$, $H_{¬\phi}$ and $p \in P$ with $H_\phi \cap H_{¬\phi} = \emptyset$, define ${\ileadsto[i^{(H_\phi, H_{¬\phi}, p)}]}$ as $\hist \ileadsto[i^{(H_\phi, H_{¬\phi}, p)}] \phi$ if $\hist \in H_\phi$, $\hist \ileadsto[i^{(H_\phi, H_{¬\phi}, p)}] ¬\phi$ if $\hist \in H_{¬\phi}$ and $\hist \ileadsto[i^{(H_\phi, H_{¬\phi}, p)}] p$ if $\hist \notin H_\phi \cup H_{¬\phi}$.

  Given any infinite set $\allargs$, define $I = \set{i^{(H_\phi, H_{¬\phi}, p)} \suchthat H_\phi \cap H_{¬\phi} = \emptyset, p \in P}$.
  Considering any $\hist \in \finite(\allhists)$ and $i^{(H_\phi, H_{¬\phi}, p)} \in I$, pick any $p' \neq p$ and consider $j = i^{(\histp, p')}$.

  NOPE

  Given any infinite set $\allargs$, define $I = \set{f \in P^{\allhists} \suchthat \pi(f) \neq \emptyset}$ and $\forall i \in I: {\ileadsto} = i$.
  Considering any $\hist \in \finite(\allhists)$ and ${\ileadsto} \in I$, pick another relation that coincides on $\hist$ but with another DJ than ${\ileadsto}$. This is possible because $\hist$ is finite, thus there is an argument that is not involved in any of its elements, which can serve as decisive one.

  SEEMS OKAY!

  $p' \neq p$ and consider $j = i^{(\histp, p')}$.

  TODO find example where Per \cref{th:caractNotIndic}, there is no claim that holds and is anonymous and indicative.
  For each path we need two individuals $i^{(\hist, \ar, p)}$ and $i^{(\hist, \ar, p')}$ that have the same behavior all along $\hist$ but different DJs.
  Defining individuals systematically in such a way should produce a suitable situation.

  If using empty DJs, it’s easier. Define $.* \ileadsto[i_1] \phi$, $\emptyset \ileadsto[i_2] \phi$, $a.* \ileadsto[i_2] \phi$, $b.* \ileadsto[i_2] ¬\phi$, $\emptyset \ileadsto[i_3] \phi$, $a.* \ileadsto[i_3] ¬\phi$, $b.* \ileadsto[i_3] \phi$. 
  Then $\pi(i_1) = \phi$ and $\pi(i_2) = \pi(i_3) = \emptyset$.
  Also, for any path, if it is only $\emptyset$, everybody agree; if it starts with $a$ then $i_1$ and $i_2$ agree on it; if it starts with $b$ then $i_1$ and $i_3$ agree on it.

  % All individuals always go to $\phi$ except for their own decisive argument, leading them, for $i_1, …$, to $\phi$, and for $j_1, …$, to $¬\phi$.
  % So that when observing someone going always to $\phi$ except for a given argument,

@@

  The condition $\pi(i_1) \neq \emptyset$ in \ref{it:condNotHAI} is required, as seen by this example where \ref{it:condNotHAI} is satisfied, even though
  $[\forall \hist \in \allhists, i_1 \in I: \exists i_2 \in I \suchthat \pi(i_1) \neq \pi(i_2) \land \forall \histp \wsub \hist: {\ileadsto[i_1]}(\histp) = {\ileadsto[i_2]}(\histp)]$ is not.
We could have \ref{it:condNotHAI} satisfied, but not the part with $\pi(i_1) \neq \emptyset$.
  Consider $i_1$ with empty DJ on some path has only individuals with identical behavior on the path that have an empty DJ (though for normal $i$’s, the condition is satisfied, thus there are collisions, meaning some individuals with different DJ having identical behavior on the path).
\end{example}
\begin{remark}
  Given $J \wsub I$, say that a set of paths discriminates $J$ iff some function associating each path to an answer is compatible with some $j \in J$ (meaning that $f \wsub {\ileadsto[j]}$) and is compatible with no $i \notin J$.
  The condition says that there must be some $p$ and some increasing set of paths $H$ such that $H$ discriminates $J = \pi^{-1}(p)$.
\end{remark}
\begin{remark}
  For failing the condition, I need an infinity of $j$’s (and at least one $i$) who agree on any branch of the history tree? (Meaning relations $\ileadsto[j]$, we don’t care about individuals.)
  No: on any finite branch in the tree, I need that remains one $j$ and one $i$ such that infinite prolongation of the branch will never discriminate them.
